---
title: "Introduction to Tree Regression with [Boston Housing Data](../datasets/boston.qmd)"
format: html
execute: 
  cache: true
---

Welcome to this tutorial on tree regression. This tutorial aims to introduce the concept of tree regression, provide a detailed explanation of the code involved, and demonstrate its implementation using R. Tree regression is a powerful tool for predicting continuous outcomes based on input features, and it is widely used in various fields.

## What is Tree Regression?

Tree regression is a type of decision tree that is used for predicting continuous target variables. The tree is constructed by recursively splitting the data into subsets based on the values of input features, aiming to minimize the variance of the target variable within each subset.

## Loading Necessary Libraries

First, we need to load the required libraries for our analysis. We will use `tidyverse` for data manipulation and visualization, `tidymodels` for model fitting, and `rpart.plot` for visualizing the tree.

```{r}
library(tidyverse)
library(tidymodels)

library(rpart.plot)
```

## Loading the Data

We load and prepare the `Boston` dataset. This dataset contains various attributes of houses in Boston, including the median value of owner-occupied homes (`medv`), which we will predict using the other attributes.

```{r load_data}
boston_data = MASS::Boston
```

## Splitting the Data

We split the data into training and testing sets using the `initial_split` function. We set a 50-50 split by setting the `prop` argument.

```{r split_data}
data_split = initial_split(boston_data, prop = 0.5)
train_set = training(data_split)
test_set = testing(data_split)
```

## Fitting the Tree Regression Model

Next, we fit a tree regression model to predict `medv` using all available predictors. We use a `recipe` to preprocess the data and define the model using `decision_tree` with specified parameters. The `workflow` function combines the recipe and model for fitting.

```{r}
preprocess_recipe = recipe(medv ~ ., data = train_set)

tree_model = decision_tree(
  cost_complexity = 0,
  tree_depth = 5,
  mode = "regression",
  min_n = 5,
  engine = "rpart"
)

tree_workflow = workflow() %>% 
  add_recipe(preprocess_recipe) %>% 
  add_model(tree_model)

tree_model_fit = tree_workflow %>% 
  fit(train_set)
```

## Making Predictions

Once the model is fitted, we use it to make predictions on the test set. The `predict` function from the workflow provides these predictions. We then evaluate the model performance using metrics such as RMSE (Root Mean Squared Error).

```{r}
predictions = tree_model_fit %>% 
  predict(test_set)

metrics = test_set %>% 
  bind_cols(predictions) %>% 
  metrics(truth = medv, estimate = .pred)

metrics
```

## Visualizing the Tree

We visualize the structure of the fitted tree to understand how the splits are made and how the predictions are generated. The `rpart.plot` function helps in creating a clear graphical representation of the tree.

```{r}
tree_model_fit %>% 
  extract_fit_engine() %>% 
  rpart.plot(roundint = FALSE)
```

## Pruning the Tree

Tree pruning is a technique used to reduce the size of a decision tree by removing sections that provide little predictive power. Pruning helps improve the model's generalizability by reducing overfitting. We use cross-validation to determine the optimal complexity parameter (`cost_complexity`) value.

We define a tree model with tunable `cost_complexity` and update the workflow to include this tunable parameter. We then define the resampling method using 5-fold cross-validation with `vfold_cv`. This function splits the data into 5 folds and performs cross-validation to evaluate the model performance.

```{r}
# Define the tree model with tunable cost complexity
tree_tune_model = decision_tree(
  cost_complexity = tune(),
  tree_depth = 5,
  mode = "regression",
  engine = "rpart",
  min_n = 5
)

# Update the workflow
tune_tree_workflow = tree_workflow %>% 
  update_model(tree_tune_model)

# Define the resampling method
set.seed(123)
cv_splits = vfold_cv(train_set, v = 5)

# Perform tuning
tune_results = tune_grid(
  tune_tree_workflow,
  resamples = cv_splits,
  grid = grid_regular(cost_complexity(range = c(-2, -0.5)), levels = 20)
)

# Extract the best parameters
best_params = select_best(tune_results, metric = "rmse")
```

The `tune_grid` function performs a grid search over a range of `cost_complexity` values to find the best parameter that minimizes the RMSE metric. We then finalize the model with the optimal `cost_complexity` parameter and fit it to the training data using `last_fit`, which refits the model on the entire training set and evaluates it on the test set.

```{r}
# Finalize the model with the best parameters
final_tree_workflow = finalize_workflow(tune_tree_workflow, best_params)

# Fit the finalized model
final_tree_fit = final_tree_workflow %>%
  last_fit(data_split)
```

We can visualize the pruned tree to compare it with the original tree.

```{r}
# Visualize the pruned tree
final_tree_fit %>% 
  extract_fit_engine() %>% 
  rpart.plot(roundint = FALSE)
```

Finally, we evaluate the performance of the pruned tree using RMSE on the test set predictions.

```{r}
final_tree_fit %>% 
  collect_metrics()
```

## Conclusion

This tutorial has provided an in-depth look at tree regression, from concept to implementation in R. By visualizing the data, fitting the model, making predictions, and visualizing the tree structure, we have demonstrated the entire process of tree regression analysis. Additionally, we have shown how to prune a decision tree using the `tidymodels` framework to improve its performance by reducing overfitting. Understanding these methods and their applications will enhance your ability to analyze and interpret continuous data effectively.
