---
title: "KNN Regression with [Boston Housing Data](../datasets/boston.qmd) - Exercise Set"
format: html
execute: 
  cache: true
---

# Exercise Set

This exercise set is designed to help you practice the steps involved in performing K-Nearest Neighbors (KNN) regression using the Boston housing dataset. Each task is followed by a chunk of code that presents the solution.

## Task 1: Load Necessary Libraries

**Task:** Load the `tidyverse` and `tidymodels` libraries.

```{r load_libraries}
#| code-fold: true
#| code-summary: "Show the code"
#| eval: false
library(tidyverse)
library(tidymodels)
```

## Task 2: Load the Data

**Task:** Select the relevant columns from the Boston housing dataset and ensure there are no duplicate rows.

```{r load_data}
#| code-fold: true
#| code-summary: "Show the code"
#| eval: false
boston_data = MASS::Boston %>% 
  select(rm, medv) %>% 
  distinct()
```

## Task 3: Fit the KNN Regression Model

**Task:** Fit a K-Nearest Neighbors (KNN) regression model to predict `medv` using `rm`. Set the number of neighbors to 100.

```{r fit_knn_regression}
#| code-fold: true
#| code-summary: "Show the code"
#| eval: false
knn_model_fit = nearest_neighbor(
  mode = "regression",
  neighbors = 100
) %>% 
  fit(medv ~ rm, data = boston_data)
```

## Task 4: Make Predictions

**Task:** Use the fitted model to make predictions.

```{r predict}
#| code-fold: true
#| code-summary: "Show the code"
#| eval: false
predictions = knn_model_fit %>% 
  predict(boston_data)
```
