---
title: "Simple Linear Regression with Palmer Penguins Data"
format: html
---

# Introduction

In this tutorial, we'll explore how to perform a simple linear regression using the Palmer Penguins dataset. Linear regression is a foundational statistical method that allows us to model the relationship between a dependent variable and one or more independent variables. Here, we'll investigate the relationship between the body mass and the bill depth of penguins.

We'll be using the `tidyverse` and `tidymodels` libraries in R for data manipulation and modeling. Let's get started!

# Load Necessary Libraries

First, we'll load the required libraries. The `tidyverse` package provides tools for data manipulation and visualization, while `tidymodels` is a suite of packages for modeling. We will also load the `palmerpenguins` package which contains the dataset.

```{r load_libraries}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

# Load the Data

Next, we'll select the relevant columns from the Palmer Penguins dataset.

```{r load_data}
penguin_data = penguins %>%
  select(species, body_mass_g, bill_depth_mm)
```

# Fit the Linear Regression Model

Now, we'll fit a simple linear regression model. Our goal is to predict `bill_depth_mm` using `body_mass_g`.

```{r fit_linear_regression}
linear_model_fit = linear_reg() %>% 
  fit(bill_depth_mm ~ body_mass_g, data = penguin_data)
```

# Inspect the Model Coefficients

After fitting the model, it's essential to inspect the coefficients to understand the relationship between the variables.

```{r coefficients}
linear_model_fit %>% 
  tidy()
```

# Make Predictions

Finally, we'll use our model to make predictions.

```{r predict}
predictions = linear_model_fit %>% 
  predict(penguin_data)
```

# Conclusion

In this tutorial, we demonstrated how to perform a simple linear regression using the Palmer Penguins dataset. We went through loading the data, fitting a model, inspecting the coefficients, and making predictions. This foundational technique can be applied to various datasets and provides a basis for more complex modeling.

Feel free to explore further by adding more variables or trying different types of models. Happy coding!
