---
title: "Simple Linear Regression with Boston Housing Data"
format: html
---

# Introduction

In this tutorial, we'll explore how to perform a simple linear regression using the Boston Housing dataset. Linear regression is a foundational statistical method that allows us to model the relationship between a dependent variable and one or more independent variables. Here, we'll investigate the relationship between the median value of owner-occupied homes (medv) and the average number of rooms per dwelling (rm).

We'll be using the `tidyverse` and `tidymodels` libraries in R for data manipulation and modeling. Let's get started!

# Load Necessary Libraries

First, we'll load the required libraries. The `tidyverse` package provides tools for data manipulation and visualization, while `tidymodels` is a suite of packages for modeling. 

```{r load_libraries}
library(tidyverse)
library(tidymodels)
```

# Load the Data

Next, we'll load the Boston Housing dataset and select the relevant columns.

```{r load_data}
boston_data = MASS::Boston %>%
  select(rm, medv)
```

# Fit the Linear Regression Model

Now, we'll fit a simple linear regression model. Our goal is to predict `medv` using `rm`.

```{r fit_linear_regression}
linear_model_fit = linear_reg() %>% 
  fit(medv ~ rm, data = boston_data)
```

# Inspect the Model Coefficients

After fitting the model, it's essential to inspect the coefficients to understand the relationship between the variables.

```{r coefficients}
linear_model_fit %>% 
  tidy()
```

# Make Predictions

Finally, we'll use our model to make predictions.

```{r predict}
predictions = linear_model_fit %>% 
  predict(boston_data)
```

# Conclusion

In this tutorial, we demonstrated how to perform a simple linear regression using the Boston Housing dataset. We went through loading the data, fitting a model, inspecting the coefficients, and making predictions. This foundational technique can be applied to various datasets and provides a basis for more complex modeling.

Feel free to explore further by adding more variables or trying different types of models. Happy coding!