---
title: "KNN Regression with Palmer Penguins Data"
format: html
---

# Introduction

In this tutorial, we'll explore how to perform K-Nearest Neighbors (KNN) regression using the Palmer Penguins dataset. KNN regression is a non-parametric method that makes predictions based on the k-nearest neighbors to a data point. This method is particularly useful when the relationship between variables is complex and nonlinear. Here, we'll investigate how bill depth is influenced by body mass using KNN regression.

We'll be using the `tidyverse` and `tidymodels` libraries in R for data manipulation and modeling. Let's get started!

# Load Necessary Libraries

First, we'll load the required libraries. The `tidyverse` package provides tools for data manipulation and visualization, while `tidymodels` is a suite of packages for modeling. We will also load the `palmerpenguins` package which contains the dataset.

```{r load_libraries}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

# Load the Data

Next, we'll select the relevant columns from the Palmer Penguins dataset. We'll remove the `species` column as it's not needed for this regression task and ensure there are no duplicate rows.

```{r load_data}
penguin_data = penguins %>% 
  select(-species) %>% 
  distinct()
```

# Fit the KNN Regression Model

Now, we'll fit a K-Nearest Neighbors (KNN) regression model. Our goal is to predict `bill_depth_mm` using `body_mass_g`. We set the number of neighbors to 100. 

The `nearest_neighbor()` function from `parsnip` is used to specify the KNN model. Here, we set the mode to "regression" and specify the number of neighbors. We then fit the model to our data.

```{r fit_knn_regression}
knn_model_fit = nearest_neighbor(
  mode = "regression",
  neighbors = 100
  ) %>%
  fit(bill_depth_mm ~ body_mass_g, data = penguin_data)
```

# Make Predictions

Finally, we'll use our model to make predictions.

```{r predict}
predictions = knn_model_fit %>% 
  predict(penguin_data)
```

# Conclusion

In this tutorial, we demonstrated how to perform KNN regression using the Palmer Penguins dataset. We went through loading the data, fitting a KNN model, and making predictions. KNN regression is a versatile technique that can capture complex relationships in data without assuming a specific functional form.

Feel free to explore further by adjusting the number of neighbors or trying different predictors. This method can be a powerful tool in your machine learning toolkit. Happy coding!
