---
title: "KNN Regression with [Boston Housing Data](../datasets/boston.qmd)"
format: html
execute: 
  cache: true
---

# Introduction

In this tutorial, we'll explore how to perform K-Nearest Neighbors (KNN) regression using the Boston housing dataset. KNN regression is a non-parametric method that makes predictions based on the k-nearest neighbors to a data point. This method is particularly useful when the relationship between variables is complex and nonlinear. Here, we'll investigate how the median value of owner-occupied homes (medv) is influenced by the average number of rooms per dwelling (rm) using KNN regression.

We'll be using the `tidyverse` and `tidymodels` libraries in R for data manipulation and modeling. Let's get started!

# Load Necessary Libraries

First, we'll load the required libraries. The `tidyverse` package provides tools for data manipulation and visualization, while `tidymodels` is a suite of packages for modeling. We will also load the `MASS` package which contains the Boston dataset.

```{r load_libraries}
library(tidyverse)
library(tidymodels)
```

# Load the Data

Next, we'll select the relevant columns from the Boston housing dataset. We'll ensure there are no duplicate rows.

```{r load_data}
boston_data = MASS::Boston %>% 
  select(rm, medv) %>% 
  distinct()
```

# Fit the KNN Regression Model

Now, we'll fit a K-Nearest Neighbors (KNN) regression model. Our goal is to predict `medv` using `rm`. We set the number of neighbors to 100. 

The `nearest_neighbor()` function from `parsnip` is used to specify the KNN model. Here, we set the mode to "regression" and specify the number of neighbors. We then fit the model to our data.

```{r fit_knn_regression}
knn_model_fit = nearest_neighbor(
  mode = "regression",
  neighbors = 100
  ) %>%
  fit(medv ~ rm, data = boston_data)
```

# Make Predictions

Finally, we'll use our model to make predictions.

```{r predict}
predictions = knn_model_fit %>% 
  predict(boston_data)
```

# Conclusion

In this tutorial, we demonstrated how to perform KNN regression using the Boston housing dataset. We went through loading the data, fitting a KNN model, and making predictions. KNN regression is a versatile technique that can capture complex relationships in data without assuming a specific functional form.

Feel free to explore further by adjusting the number of neighbors or trying different predictors. This method can be a powerful tool in your machine learning toolkit. Happy coding!
