---
title: "Introduction to the Boston Data Set"
format: html
---

The Boston data set, famously used in various machine learning and statistical analysis projects, provides a comprehensive look at housing values in the suburbs of Boston. It has been widely used for regression tasks, particularly predicting house prices based on various features. This data set, which originates from the U.S. Census Bureau, contains 506 observations of 14 different variables, each representing different characteristics of the suburbs in Boston. 


## Variables in the Boston Data Set

Here is an explanation of each variable in the Boston data set, along with their names and data types:
  
  1. **crim**: Per capita crime rate by town (Numeric)
  2. **zn**: Proportion of residential land zoned for lots over 25,000 sq. ft.        (Numeric)
  3. **indus**: Proportion of non-retail business acres per town (Numeric)
  4. **chas**: Charles River dummy variable (1 if tract bounds river; 0 otherwise)     (Binary)
  5. **nox**: Nitrogen oxides concentration (parts per 10 million) (Numeric)
  6. **rm**: Average number of rooms per dwelling (Numeric)
  7. **age**: Proportion of owner-occupied units built prior to 1940 (Numeric)
  8. **dis**: Weighted distances to five Boston employment centers (Numeric)
  9. **rad**: Index of accessibility to radial highways (Ordinal)
  10. **tax**: Full-value property tax rate per $10,000 (Numeric)
  11. **ptratio**: Pupil-teacher ratio by town (Numeric)
  12. **black**: 1000(Bk - 0.63)^2 where Bk is the proportion of Black residents    by town (Numeric)
  13. **lstat**: Percentage of lower status of the population (Numeric)
  14. **medv**: Median value of owner-occupied homes in $1000s (Numeric)

## Loading the Boston Data Set

In R, data sets from specific packages are often loaded using the `::` operator, which allows you to access a dataset or function from a particular package without attaching the entire package. This is particularly useful when you want to avoid name conflicts or simply reduce the clutter in your namespace.

To load the Boston data set from the `MASS` package, you can use the following code:
  
```{r }
#| eval: false

# Load the Boston data set
data("Boston", package = "MASS")

```

Here’s a step-by-step explanation of the code:
  
  1. **data("Boston", package = "MASS")**: This line loads the Boston data set. The `data` function is a base R function that loads specified data sets. By providing the name of the data set (`"Boston"`) and the package from which it comes (`"MASS"`), R understands where to find the data set.

2. **MASS::Boston**: Alternatively, you can directly reference the data set using the `::` operator as follows:
  
```{r }

Boston <- MASS::Boston

```

The `::` operator allows you to access the `Boston` data set from the `MASS` package directly. This method does not load the entire `MASS` package into your R session, but only the specific data set or function you are interested in.

