---
title: "Introduction to the Default Data Set"
format: html
---

The Default data set, included in the ISLR package, is frequently utilized in classification tasks within machine learning and statistical analysis. This data set comprises information about 10,000 individuals, offering insights into factors that might influence loan default behavior. It includes variables indicating default status, student status, average credit card balance, and income. 

## Variables in the Default Data Set

Here is an explanation of each variable in the Default data set, along with their names and data types:
  
  1. **default**: Indicates whether the individual defaulted on their loan (Factor: `No`, `Yes`)
  2. **student**: Indicates whether the individual is a student (Factor: `No`, `Yes`)
  3. **balance**: The average balance that the individual has remaining on their credit card (Numeric)
  4. **income**: The individual's income (Numeric)

## Loading the Default Data Set

In R, data sets from specific packages are often loaded using the `::` operator, which allows you to access a dataset or function from a particular package without attaching the entire package. This is particularly useful to avoid name conflicts or reduce the clutter in your namespace.

To load the Default data set from the `ISLR` package, you can use the following code:
  
```{r}
#| eval: false

# Load the Default data set
data("Default", package = "ISLR")
```

Here’s a step-by-step explanation of the code:
  
1. **data("Default", package = "ISLR")**: This line loads the Default data set. The `data` function is a base R function that loads specified data sets. By providing the name of the data set (`"Default"`) and the package from which it comes (`"ISLR"`), R understands where to find the data set.

2. **ISLR::Default**: Alternatively, you can directly reference the data set using the `::` operator as follows:
  
```{r}
Default = ISLR::Default
```

The `::` operator allows you to access the `Default` data set from the `ISLR` package directly. This method does not load the entire `ISLR` package into your R session, but only the specific data set or function you are interested in.